Finite Property Convex Elicitation Notes

### 2018-05-08 cvx-fin-notes.tex
more general d stuff

\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

@@

\begin{lemma}
  Let $L$ be a convex loss function.
  Then $\Gamma_L$ is non-degenerate if and only if $L(\cdot,y)$ is bounded from below for all $y\in\Y$.
\end{lemma}

\begin{definition}
  We say a loss $L$ is \emph{finite-piecewise-linear and convex (FPLC)} if for all $y\in\Y$, the function $L(\cdot)_y$ is a piecewise-linear function with finitely many pieces, and is bounded from below.
\end{definition}

@@

\begin{lemma}\label{lem:fplc-directional-deriv}
  Let $L$ be FPLC, and let $r\in\reals$ be a point of differentiability of $L$.
  Let $d^-(r)_y$ and $d^+(r)_y$ denote the left and right derivative, respectively, of $L(\cdot)_y$ at $r$, so that $d^-(r),d^+(r)\in\reals^\Y$.
  Then for any open interval $(a,b)$ containing $r$ where $L$ is differentiable, we have
$d^+(a) = d^-(r) = d^+(r) = d^-(b)$.
\end{lemma}

\begin{lemma}\label{lem:simple-intervals}
  Let $a<a'$ and $b\leq b'$ such that either $a< 0 < b$ or $b' < 0 < a'$.
  Then there exists $\alpha \in (0,1)$ such that $0 \in \alpha [b,b'] + (1-\alpha) [a,a']$.
\end{lemma}
\begin{proof}
  Suppose $a < 0 < b$.
  Let $w = (a + \min(a',0))/2 < 0$ and take $\alpha = w/(w-b) \in (0,1)$.
  Then $\alpha b + (1-\alpha) w = wb/(w-b) - bw/(w-b) = 0$.
  As $a < w < a'$, the result follows:
  \begin{align*}
    & \alpha b + (1-\alpha) a < \alpha b + (1-\alpha) w = 0~,
    \\
    & \alpha b' + (1-\alpha) a' > \alpha b + (1-\alpha) w = 0~.
  \end{align*}
  Similarly, if $b' < 0 < a'$, we take $w = (\max(a,0)+a')/2 > 0$, and same construction holds.
\end{proof}

@@

@@

\raf{I further conjecture that there is a convex subset $C$ of $\reals^d$ such that $\R' = \R\cap C$ gives $\trim(\Gamma_L) = \Gamma_L\cap \R'$, just like in the 1-d case.  If that's true, it would be really helpful in figuring out what the constraints are in the simplex between neighbering reports!}

>>>>

\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}

@@

\begin{lemma}
  Let $L$ be a convex loss function.
  Then $\Gamma_L$ is non-degenerate if and only if $L(\cdot,y)$ is bounded from below for all $y\in\Y$.
  \raf{Not true; take $L(r,y) = e^r$}
\end{lemma}

\begin{definition}
  We say a loss $L:\reals^d \to \reals^\Y$ is \emph{finite-piecewise-linear and convex (FPLC)} if for all $y\in\Y$, the function $L(\cdot)_y$ is a piecewise-linear function with finitely many pieces, convex, and bounded from below.
\end{definition}

@@

\begin{lemma}\label{lem:fplc-directional-deriv}
  Let $L:\reals^d \to \reals^\Y$ be FPLC, and let $r\in\reals$ be a point of differentiability of $L$.
  Let $d^-(r)_y$ and $d^+(r)_y$ denote the left and right derivative, respectively, of $L(\cdot)_y$ at $r$, so that $d^-(r),d^+(r)\in\reals^\Y$.
  Then for any open interval $(a,b)$ containing $r$ where $L$ is differentiable, we have
$d^+(a) = d^-(r) = d^+(r) = d^-(b)$.
\end{lemma}

\begin{lemma}\label{lem:simple-intervals}
  Let $a<a'$ and $b\leq b'$ such that either $a< 0 < b$ or $b' < 0 < a'$.
  Then there exists $\alpha \in (0,1)$ such that $\alpha b + (1-\alpha) a < 0 < \alpha b' + (1-\alpha) a'$.
\end{lemma}
\begin{proof}
  Suppose $a < 0 < b$.
  Let $w = (a + \min(a',0))/2 < 0$ and take $\alpha = w/(w-b) \in (0,1)$.
  Then $\alpha b + (1-\alpha) w = wb/(w-b) - bw/(w-b) = 0$.
  As $a < w < a'$, the result follows:
  \begin{align*}
    & \alpha b + (1-\alpha) a < \alpha b + (1-\alpha) w = 0~,
    \\
    & \alpha b' + (1-\alpha) a' > \alpha b + (1-\alpha) w = 0~.
  \end{align*}
  Similarly, if $b' < 0 < a'$, we take $w = (\max(a,0)+a')/2 > 0$, and same construction holds.
\end{proof}

@@

\begin{definition}
  Let $f:\reals^d\to\reals$ be FPLC.
  A \emph{vertex} of $f$ is a vertex of the corresponding convex polytope.
\end{definition}

\begin{lemma}\label{lem:vertiex-subgrad}
  Let $f:\reals^d\to\reals$ be FPLC.
  For every non-vertex $x$ of $f$, there is a vertex $v$ such that $\partial f(x) \subseteq \partial f(v)$.
\end{lemma}
\begin{proof}
  \raf{I'm thinking of a proof in two parts.  1. lemma from convex polytopes saying that you can always get to a vertex while staying within the current face; 2. all points in a face but no other face (so the relative interior) share the same subgradient set, and at the relative boundary, a superset of that set.}
\end{proof}

@@

\raf{If true, this will be helpful in figuring out what the constraints are in the simplex between neighbering reports!}

Note that the definition of $\R'$ is a generalization of its definition as $[\underline r, \overline r]$ in Proposition~\ref{prop:fplc-trim-char}.
\begin{conjecture}
  Let $L:\reals^d \to\reals^\Y$ be FPLC, and $\R$ the vertices of $\sum_{y\in\Y} L(\cdot)_y$.
  Then $\trim(\Gamma_L) \equiv \Gamma_L\cap \R'$ where $\R' = \conv \{ \argmin_{r\in\reals^d} L(r)_y \cap \R : y\in\Y\}$.
\end{conjecture}

\begin{conjecture}
  Let $L:\reals^d \to\reals^\Y$ be FPLC, and $\R$ the vertices of $f(\cdot) = \sum_{y\in\Y} L(\cdot)_y$.
  Then for all $p\in\inter\Delta(\Y)$, the convex hull of $\Gamma_L(p)\cap\R$ is a face of the polytope corresponding to $f$.
\end{conjecture}
\begin{proof}
  \raf{Sketch:}
  The polytope of $f$ is the same as that of $p\cdot L(\cdot)$, since $p$ has full support.  As $p\cdot L$ must be flat between all of $\Gamma_L(p)$, those vertices are on the same face.  Moreover, all vertices on the minimal face containing $\Gamma_L(p)$ are optimal.
\end{proof}
